# 4.4 For-loops

```{r}
compute_s_n <- function(n){
 x <- 1:n
 sum(x)
}
```
simplest loop

```{r}
for(i in 1:5){
 print(i)
}
```
Sn Example

```{r}
m <- 25
s_n <- vector(length = m)
for(n in 1:m){
 s_n[n] <- compute_s_n(n)
}
```

```{r}
n <-1:m
plot(n, s_n)
```

confirm with a table
```{r}
head(data.frame(s_n = s_n, formula = n*(n+1)/2))
```

```{r}
plot(n, s_n)
lines(n, n*(n+1)/2)
```

# 4.5 Vectorization and Functionals
Vectorized function
```{r}
x <- 1:10
sqrt(x)
y <- 1:10
x*y
```

```{r}
n <- 1:25
compute_s_n(n)
```
sapply permits us to perform element-wise operations on any function.
```{r}
x <- 1:10
sapply(x, sqrt)
```
```{r}
n <- 1:25
s_n <- sapply(n, compute_s_n)
plot(n, s_n)
```

# 4.6 Exercises

1.
What will this conditional expression return?
```{r}
x <- c(1,2,-3,4)

if(all(x>0)){
  print("all positives")
} else{
  print("Not all positives")
}
```
A. Not all posictives.

2.
Which of the following expressions is always FALSE when at least one entry of a logical vector x is TRUE?
```{r}
x <- c(TRUE, FALSE, TRUE)
all(x)
any(x)
any(!x)
all(!x)
```
D.

3.
The function nchar tells you how many characters long a character vector is.
Write a line of code that assigns to the object new_names the state abbreviation when the state name is longer than 8 characters.

```{r}
library(dslabs)
data(murders)
new_names <- ifelse(nchar(murders$state)>8, murders$abb, murders$state)
new_names
```

4.
Create a function sum_n that for any given value, say n, computes the sum of the integers from 1 to n (inclusive). Use the function to determine the sum of integers from 1 to 5,000.
```{r}
sum_n <- function(n){
  x <- 1:n
  sum(x)
}
sum_n(5000)
```

5.
Create a function altman_plot that takes two arguments, x and y, and plots the difference against the sum.
```{r}
altman_plot <- function(x, y){
  plot(x+y, y-x)
}
```

6.
After running the code below, what is the value of x?
```{r}
x <- 3
my_func <- function(y){
  x <- 5
  y+5

}
x
```
3

7.
```{r}
compute_s_n <-function(n){
  x <- 1:n
  sum(x*x)
}
compute_s_n(10)
```
385

8. 
```{r}
s_n <- vector("numeric", 25)
n <- 1:25
for(i in n){
  s_n[i] <- compute_s_n(i)
}
s_n
```

9.
```{r}
s_n <- sapply(n, compute_s_n)
s_n
```

10.
```{r}
s_n <- map_dbl(n, compute_s_n)
s_n
```

11.
```{r}
n <- 1:25
plot(n, s_n)
```

12.
```{r}
n <- 1:25
plot(n, s_n)
lines(n, n*(n+1)*(2*n+1)/6)
```
